Stateless Model Checking with Data-Race Preemption Points

### 2015-11-17 pldi-16/evaluation.tex
Fix the story about partial verification. Fill out the table, mostly.

\newcommand\mxtest{\texttt{mx\_test}}
\newcommand\tej{\texttt{join\_test}}
\newcommand\bct{\texttt{bcast\_test}}
\newcommand\paraguay{\texttt{signal\_test}}
\newcommand\paradise{\texttt{sem\_test}}
\newcommand\rwldgr{\texttt{rwlock\_test}}
We tested P2s with 6 multithreaded programs:
% from the 410 test suite % XXX: I would like to say this but this is a lie; figure out what else i can say instead
% each tailored to exercise a different part of the P2 project
\mxtest, for locking algorithm correctness,
\tej, a test of thread lifecycle,
\bct~and \paraguay~for cvars,
\paradise~for semaphores,
and \rwldgr~for r/w locks.
For \mxtest, \paradise, and \paraguay, we used {\tt without\_function} to blacklist {\tt thr\_create}, {\tt thr\_exit}, and {\tt thr\_join},
and for \mxtest~we enabled \landslide's mutex-testing option
(see \sect{\ref{sec:landslide}}).
\newcommand\prisema{\texttt{sched\_test}}
\newcommand\waitsimple{\texttt{wait\_test}}
\newcommand\alarmsimul{\texttt{alarm\_test}}
We tested the Pintoses with 3 programs from the class test suite: \prisema, a test of the kernel scheduling algorithm,
\alarmsimul, for the timer sleep routine,
and \waitsimple, for process lifecycle system calls\footnote{
	Some of the Pintoses were partially implemented, so each test program could only be run on a subset of the 78 submissions; see Table~\ref{tab:drbugs}.
}
The source code of all 9 test cases is available at
{\em [submitted as supplementary material]}.
For all tests, we used {\tt without\_function} to blacklist PPs on the {\tt malloc} lock.
We ran all tests on 12-core 3.2 GHz Xeon machines with 12GB of RAM.
In total, our evaluation comprises 629 tests (pairs of a test program and a Pintos or P2),
at least 163 of which could expose bugs.

@@

\begin{table*}[t]
	\begin{center}
		\small
	\begin{tabular}{r|c||c|c|c||c|c|c|c|c}
		& {\bf Num.} & {\bf SSS-MC} & {\bf QS} & {\bf DR-only} & {\bf Total} & {\bf Untested} & {\bf Malloc-} & {\bf Nondet.} & {\bf Nondet.} \\
		{\bf Test name} & {\bf tested} & {\bf bugs} & {\bf bugs} & {\bf bugs} & {\bf DR PPs} & {\bf DR PPs} & {\bf recycle DRs} & {\bf DR PPs} & {\bf DR bugs} \\
		\hline
		{\bct		} & 79	& 7	& 11	& 4	&	&	& 65	&	& 4	\\
		{\tej		} & 79	& 14 	& 25	& 11	&	&	& 471	&	& 6	\\
		{\mxtest	} & 79	& 1	& 10	& 9	&	&	& 7	&	& 1	\\
		{\paradise	} & 79	& 10	& 17	& 7	&	&	& 268	&	& 3	\\
		{\paraguay	} & 79	& 4	& 12	& 9	&	&	& 188	&	& 0	\\
		{\rwldgr	} & 79	& 26	& 30	& 3	&	&	& 157	&	& 3	\\
		\hline
		{\prisema	} & 59	& \multicolumn{5}{c}{TODO: finish this table} & 0	&	&	\\
		{\alarmsimul	} & 44	& \multicolumn{5}{c}{haven't yet decided how to organize this data;} & 37	&	&	\\
		{\waitsimple	} & 52	& \multicolumn{5}{c}{suggestions welcome} & 57	&	&	\\
		\hline
		{\bf Total}	& 629	&	&	&	&	&	& 1250	&	&
	\end{tabular}
	\end{center}
	\caption{Summary of number and types of bugs found by each test program.
		SSS-MC is the control; QS is \quicksand.
		``DR-only'' counts among \quicksand's bugs how many required data-race PPs to expose (\sect{\ref{sec:eval-sssmc}}).
	``Nondeterministic DRs'' counts among the DR-only bugs how many candidates required MC integration to identify (\sect{\ref{sec:eval-falseneg}}).
		TODO: more explanation}
	\label{tab:drbugs}
\end{table*}

@@

{\bf Partial verification.}
When a state space times out, the user may prefer a brief summary of what parts of the test were verified, rather than writing off all the CPU time as a waste.
Beyond using the state space estimator to guess the percent coverage,
we know of no approach to quantify the probability
that a bug would be exposed by an untested interleaving.
\quicksand~does not attempt any such quantitative guarantee,
but can at least report which subsets of PPs produced state spaces that did complete in time.

On 77 tests, the SSS-MC control experiment timed out in 10 hours with no bugs found.
Among these, \quicksand~found data-race bugs in 15.
For the other 62, we counted the number of state spaces \quicksand~was able to complete and the number of unique PPs among them.
In three cases, \quicksand~also failed to complete anything; beyond these,
between 1 and 253 state spaces were completed, with a mean of 58.9 and median of 29.
Between 0 and 233 unique PPs were tested among some completed subsets, with a mean of 27.2 and median of 7.5.
These completions guarantee that, if the test program could expose a bug,
% Justify implicit hypothesis: Sync PPs + DR PPs = all possible relevant PPs.
it would only be found by a new data-race PP not discovered yet, or by a superset combination of PPs not reached.

@@

Among the 69 true data-races which initially fit the malloc-recycle pattern,
none of them exposed a new bug when used as a PP.
This shows that for other data-race tools,
suppressing malloc-recycle candidates may be a productive heuristic,
even if unsound without Iterative Deepening.
However, \quicksand~was able to correctly identify the 69 violations of that heuristic (among 30 distinct tests),
and fall back to classifying them with DPOR.
%which we consider a much stronger verification.

>>>>

\newcommand\mxtest{\texttt{mx\_test}}
\newcommand\tej{\texttt{join\_test}}
\newcommand\bct{\texttt{bcast\_test}}
\newcommand\paraguay{\texttt{signal\_test}}
\newcommand\paradise{\texttt{sem\_test}}
\newcommand\rwldgr{\texttt{rwlock\_test}}
We tested P2s with 6 multithreaded programs:
% from the 410 test suite % XXX: I would like to say this but this is a lie; figure out what else i can say instead
% each tailored to exercise a different part of the P2 project
\mxtest, for locking algorithm correctness,
\tej, a test of thread lifecycle,
\bct~and \paraguay~for cvars,
\paradise~for semaphores,
and \rwldgr~for r/w locks.
For \mxtest, \paradise, and \paraguay, we used {\tt without\_function} to blacklist {\tt thr\_create}, {\tt thr\_exit}, and {\tt thr\_join},
and for \mxtest~we enabled \landslide's mutex-testing option
(see \sect{\ref{sec:landslide}}).
\newcommand\prisema{\texttt{sched\_test}}
\newcommand\waitsimple{\texttt{wait\_test}}
\newcommand\alarmsimul{\texttt{alarm\_test}}
We tested the Pintoses with 3 programs from the class test suite: \prisema, a test of the kernel scheduling algorithm,
\alarmsimul, for the timer sleep routine,
and \waitsimple, for process lifecycle system calls\footnote{
	Some of the Pintoses were partially implemented,
	so each test could only be run on a subset of the 78 submissions; see ``Number tested'' in Table~\ref{tab:drbugs}.
}
The source code of all 9 test cases is available at
{\em [submitted as supplementary material]}.
For all tests, we used {\tt without\_function} to blacklist PPs on the {\tt malloc} lock.
We ran all tests on 12-core 3.2 GHz Xeon machines with 12GB of RAM.
In total, our evaluation comprises 629 tests (pairs of a test program and a Pintos or P2),
at least 163 of which could expose bugs.

@@

\begin{table*}[t]
	\begin{center}
		\small
	\begin{tabular}{r|c||c|c|c|c|c||c|c||c|c}
		& {\bf Num.} & {\bf SSS-MC} & {\bf QS} & {\bf DR-only} & {\bf Nondet.} & {\bf Malloc-} &
		{\bf Mutual} & {\bf Avg. tested} & {\bf Total} & {\bf Untested} \\
		{\bf Test name} & {\bf tested} & {\bf bugs} & {\bf bugs} & {\bf bugs} & {\bf DR bugs} & {\bf recycle DRs} &
		{\bf time-outs} & {\bf subset SSes} & {\bf DR PPs} & {\bf DR PPs} \\
		\hline
				% #test  sssmc   qs      dronly  nondets  FRMs   timeout comp.SSes DRPPs untested DRPPs
		{\tt bcast}	& 79	& 7	& 11	& 4	& 4	& 65	& 0	& -	&	&	\\
		{\tt join} 	& 79	& 14 	& 25	& 10	& 6	& 471	& 5	& 60.8	&	&	\\
		{\tt mx} 	& 79	& 1	& 10	& 9	& 1	& 7	& 0	& -	&	&	\\
		{\tt sem} 	& 79	& 10	& 17	& 6	& 5	& 268	& 29	& 73.3	&	&	\\
		{\tt signal} 	& 79	& 4	& 12	& 7	& 6	& 188	& 10	& 54.2	&	&	\\
		{\tt rwlock} 	& 79	& 26	& 30	& 2	& 2	& 157	& 19	& 29.6	&	&	\\
		\hline
		{\tt sched} 	& 59	& 1	& 7	& 6	& 4	& 0	& 0	& -	&	&	\\
		{\tt alarm} 	& 44	& 5	& 21	& 17	& 3	& 37	& 22	& 9.1	&	&	\\
		{\tt wait} 	& 52	& 23	& 30	& 7	& 1	& 57	& 8	& 28.0	&	&	\\
		\hline
		{\bf Total}	& 629	& 91	& 163	& 68	& 32	& 1250	& 93	& 42.6 	&	&	\\
	\end{tabular}
	\end{center}
	\caption{Summary of bugs and data races found by each test program.
		SSS-MC is the control; QS is \quicksand.
		``DR-only bugs'' counts among \quicksand's bugs how many required data-race PPs to expose (\sect{\ref{sec:eval-sssmc}});
	among those, ``Nondeterministic DR bugs'' counts how many candidates required MC integration to identify (\sect{\ref{sec:eval-falseneg}}).
	``Malloc-recycle DRs'' counts how many false positives we suppressed (\sect{\ref{sec:eval-falsepos}}).
	``Mutual time-outs'' counts how often both SSS-MC and \quicksand~timed out with no bug found;
	among those, ``Average tested subset SSes'' counts how many partial verifications \quicksand~provided on average for each test (\sect{\ref{sec:eval-sssmc}}).
	``Total DR PPs'' counts how many unique data-racing instructions we identified among tests with no bugs found;
	among those, ``Untested DR PPs'' counts how many could not be checked in the time limit (\sect{\ref{sec:future}}).
		}
	\label{tab:drbugs}
\end{table*}

@@

{\bf Partial verification.}
When a state space times out, the user may prefer a brief summary of what parts of the test were verified, rather than writing off all the CPU time as a waste.
%Beyond using the state space estimator to guess the percent coverage,
We know of no approach to quantify the probability
that a bug would be exposed by an untested interleaving,
but \quicksand~at least reports which subsets of PPs produced state spaces that did complete in time.
%\quicksand~does not attempt any such quantitative guarantee,
%but can at least report
On 129 tests, the SSS-MC control experiment timed out in 10 hours with no bugs found.
Among these, \quicksand~found bugs in 36.
For the other 93, we show the number of state spaces \quicksand~was able to complete in the ``Average tested subset SSes'' column of Table~\ref{tab:drbugs}.
%and the number of unique PPs among them.
%In 6 cases, \quicksand~also failed to complete anything; beyond these,
%between 1 and 253 state spaces were completed for each test
%Between 0 and 233 unique PPs were tested among some completed subsets, with a mean of 27.2 and median of 7.5.
These completions guarantee that, if the test program could expose a bug,
% Justify implicit hypothesis: Sync PPs + DR PPs = all possible relevant PPs.
it would only be found by a new data-race PP not discovered yet, or by a superset combination of PPs not reached.

@@

Among those 69 true data-races, %which initially fit the malloc-recycle pattern,
none of them exposed a new bug when used as a PP.
This shows that for other data-race tools,
suppressing malloc-recycle candidates may be a productive heuristic,
even if unsound without Iterative Deepening.
However, \quicksand~was able to correctly identify the 69 violations of that heuristic (among 30 distinct tests),
and fall back to classifying them with DPOR.
%which we consider a much stronger verification.

### 2015-11-17 pldi-16/paper.tex
begin major swizzling around of sections according to joan's suggestion
Write overview section
TRIMTRIMTRIM. rip in pieces, author-year citations :(
split discussion into own tex file
Fix the story about partial verification. Fill out the table, mostly.

% TODO: Don't forget to run e.g. s/\\quicksand /\\quicksand~/
\newcommand\landslide{\textsc{Landslide}}
\newcommand\quicksand{\textsc{Quicksand}}
\newcommand\simics{\textsc{Simics}}
\newcommand{\sect}[1]{\S #1}
\newcommand\hilight[2]{\color{#1}#2\color{black}}
\definecolor{olivegreen}{RGB}{0,127,0}
\definecolor{brickred}{RGB}{192,0,0}
\definecolor{commentblue}{RGB}{0,0,255}

@@

We present an {\em Iterative Deepening} framework for stateless model checking,
which manages the exploration of many state spaces using different preemption points.
It uses state space estimation to prioritize jobs most likely to complete in a fixed CPU budget,
and it incorporates data-race analysis to add new preemption points on the fly.
Preempting threads during a data race's instructions
%These preemption points allow us to check whether each data-race can produce an observable failure, and
can automatically classify the race as buggy or benign,
and uncovers new bugs not reachable by prior model checkers.
%
It also enables full verification of all possible schedules when every data race is verified as benign within the CPU budget.
In our evaluation, Iterative Deepening
%with data-race preemption points
found 79\% more bugs than the best prior stateless model checking approach.
%Our evaluation shows this technique is
%more effective than single-state-space model checking
%%both at finding more bugs and at completing more state spaces when no bug exists.
%both at finding the same bugs faster and at finding new bugs entirely.

@@

\input{intro}
\input{di}
\input{evaluation}
\input{related}

@@

We have presented Iterative Deepening and \quicksand, a new technique and tool for automating the choice of PPs during stateless model checking.
% , and \quicksand, a tool which implements this technique tailored to the \landslide~model checker.
\quicksand~incorporates data-race analysis to create new PPs tailored specifically to the program under test,
and manages multiple model checker instances to test many different PP subsets in a given CPU budget, even when the full state space of all PPs would be computationally intractable.

Our evaluation shows that \quicksand~achieves better bug-finding results than either single-pass data-race detection or single-state-space model checking (SSS-MC) alone.
%We dramatically reduce false-positive data race reports,
%%by verifying the absence of bugs in their associated state spaces
%and find bugs caused by nondeterministic data-races in alternate thread interleavings missed entirely by single-pass analysis.
Compared to SSS-MC,
%, we find bugs faster in smaller subset state spaces, %% not true, or at least, not by very much
%when the maximal state space is too large to explore
we find new bugs with data-race PPs that the maximal state space would not expose at all.
Moreover, when all data-race PPs can be fully tested within the CPU budget, we provide an equally strong verification as if every single instruction had been used as a PP.

>>>>

% TODO: Don't forget to run e.g. s/\\quicksand /\\quicksand~/
\newcommand\landslide{\textsc{Landslide}}
\newcommand\quicksand{\textsc{Quicksand}}
\newcommand\simics{\textsc{Simics}}
\newcommand{\sect}[1]{\S #1}
\newcommand\hilight[2]{\color{#1}#2\color{black}}
\definecolor{orange}{RGB}{192,96,0}
\definecolor{olivegreen}{RGB}{0,127,0}
\definecolor{brickred}{RGB}{192,0,0}
\definecolor{commentblue}{RGB}{0,0,192}

@@

We present an {\em Iterative Deepening} framework for stateless model checking,
which manages the exploration of many state spaces using different preemption points.
It uses state space estimation to prioritize jobs most likely to complete in a fixed CPU budget,
and it incorporates data-race analysis to add new preemption points on the fly.
Preempting threads during a data race's instructions
%These preemption points allow us to check whether each data-race can produce an observable failure, and
can automatically classify the race as buggy or benign,
and uncovers new bugs not reachable by prior model checkers.
%
It also enables full verification of all possible schedules when every data race is verified as benign within the CPU budget.
In our evaluation, Iterative Deepening
%with data-race preemption points
found 1.79x as many bugs and verified 4.25x as many tests compared to stateless model checking without data-race preemption points.
%Our evaluation shows this technique is
%more effective than single-state-space model checking
%%both at finding more bugs and at completing more state spaces when no bug exists.
%both at finding the same bugs faster and at finding new bugs entirely.

@@

\input{intro}
\input{overview}
\input{design}
\input{implementation}
\input{soundness}
\input{evaluation}
\input{warpzone}
\input{related}

@@

We have presented Iterative Deepening and \quicksand, a new technique and tool for automating the choice of preemption points (PPs) during stateless model checking.
% , and \quicksand, a tool which implements this technique tailored to the \landslide~model checker.
\quicksand~incorporates data-race analysis to create new PPs tailored specifically to the program under test,
and manages multiple model checker instances to test many different PP subsets in a given CPU budget, even when the full state space of all PPs would be computationally intractable.

Our evaluation shows that \quicksand~achieves better bug-finding results than either single-pass data-race detection or single-state-space model checking alone.
%We dramatically reduce false-positive data race reports,
%%by verifying the absence of bugs in their associated state spaces
%and find bugs caused by nondeterministic data-races in alternate thread interleavings missed entirely by single-pass analysis.
%Compared to prior work,
%, we find bugs faster in smaller subset state spaces, %% not true, or at least, not by very much
%when the maximal state space is too large to explore
We find new bugs with data-race PPs that could not be exposed by preempting only on synchronization APIs.
Moreover, when all data-race PPs can be fully tested within the CPU budget, we provide an equally strong verification as if every single instruction had been used as a PP.

### 2015-11-17 pldi-16/warpzone.tex
Fix the story about partial verification. Fill out the table, mostly.

{\bf Finer-grained PP subsets.}
% TODO: update these numbers
\quicksand~was able to partially guarantee safety in 97\% of tests with large maximal state spaces.
However, in 2 tests, no more than the minimal state space could be verified,
and in 3 other tests, no state spaces could be completed at all.
%Larger state spaces often result from finer-grained locking,
%which can indicate a more intricate concurrent algorithm or an unnecessarily complicated design.
%Such programs may require even more rigorous verification than a program with a single global lock.
%Hence these corner cases are important to consider for future work.
While we used {\tt within\_function} (\sect{\ref{sec:landslide}}) {\em statically} to restrict where PPs could arise in advance of the test,
%we envision
future
%Iterative Deepening
implementations could use this mechanism to {\em dynamically} subset PPs further,
making partial verification of such large tests possible.
%enabling partial verification of such large tests. % if need space
%Because our current implementation does not avoid repeated interleavings across state spaces,
%as discussed above,
%we were limited to a small number of very basic PP subsets to statically seed the exploration

@@

{\bf Partial verification.}
%We are not the first to provide a partial verification guarantee when timing out on too-large state spaces (\sect{\ref{sec:eval-sssmc}}).
While we guarantee safety when preempted on certain combinations of PPs (\sect{\ref{sec:eval-sssmc}}),
CHESS
guarantees safety under no more than a certain number of preemptions \cite{chess-icb}.
%according to the maximum bound reached in the time limit.
%We imagine
These two guarantees could be each be useful to developers in different scenarios,
and future work could combine the two approaches to provide both at once.
One benefit of our technique is that {\tt within\_function} %-based Iterative Deepening (discussed above)
would enable expert developers to
%configure custom subsets of PPs they are most interested in verifying,
%according to which modules of a codebase they wish to test.
restrict Iterative Deepening to only the modules of a codebase they wish to test.

Likewise, when full verification is not computationally feasible,
some jobs with data-race PPs will time out.
We cannot guarantee that those races are
%false positives or
benign, even though no bug was found.
% FIXME: You need a separate table counting only the state spaces WITHOUT BUGES.
In the ``Untested DR PPs'' column of Table~\ref{tab:drbugs}, we count the number of such unverified candidates from each test program.
For a more formal treatment of these cases, we refer the reader to the {\em k-witness harmless} metric introduced by \cite{portend},
which could be combined with Iterative Deepening in future work.
% Future work: Add parallel DPOR so you can fill your spare CPUs when there are fewer than the max number of jobs.

>>>>

{\bf Finer-grained PP subsets.}
\quicksand~was able to partially guarantee safety for some PPs in 89\% of tests with too-large maximal state spaces.
However, in 4 tests, no more than the minimal state space could be verified,
and in 6 other tests, no state spaces were completed at all.
%Larger state spaces often result from finer-grained locking,
%which can indicate a more intricate concurrent algorithm or an unnecessarily complicated design.
%Such programs may require even more rigorous verification than a program with a single global lock.
%Hence these corner cases are important to consider for future work.
While we used {\tt within\_function} (\sect{\ref{sec:landslide}}) {\em statically} to restrict where PPs could arise in advance of the test,
%we envision
future
%Iterative Deepening
implementations could use this mechanism to {\em dynamically} subset PPs further,
making partial verification of such large tests possible.
%enabling partial verification of such large tests. % if need space
%Because our current implementation does not avoid repeated interleavings across state spaces,
%as discussed above,
%we were limited to a small number of very basic PP subsets to statically seed the exploration

@@

{\bf Partial verification.}
%We are not the first to provide a partial verification guarantee when timing out on too-large state spaces (\sect{\ref{sec:eval-sssmc}}).
While we guarantee safety when using certain combinations of PPs (\sect{\ref{sec:eval-sssmc}}),
Iterative Context Bounding
guarantees safety under no more than a certain number of preemptions \cite{chess-icb}.
%according to the maximum bound reached in the time limit.
%We imagine
These guarantees could each be useful to developers in different scenarios,
and future work could combine the two approaches to provide both at once.
One benefit of our technique is that {\tt within\_function} %-based Iterative Deepening (discussed above)
would enable expert developers to
%configure custom subsets of PPs they are most interested in verifying,
%according to which modules of a codebase they wish to test.
restrict Iterative Deepening to only the modules of a codebase they wish to test.

Likewise, when full verification is not computationally feasible,
some jobs with data-race PPs will time out.
We cannot guarantee that those races are
%false positives or
benign, even though no bug was found.
% FIXME: You need a separate table counting only the state spaces WITHOUT BUGES.
In the ``Untested DR PPs'' column of Table~\ref{tab:drbugs}, we show how many such candidates we could not verify.
For a more formal treatment of these cases, we refer the reader to the {\em k-witness harmless} metric introduced by \cite{portend},
which could be combined with Iterative Deepening in future work.
% Future work: Add parallel DPOR so you can fill your spare CPUs when there are fewer than the max number of jobs.
